Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-05-29 manuscript.tex
Bring tex up to speed

Using cross-recurrence quantification analysis (CRQA; Zbilut et al., 1998)
and windowed cross-recurrence quantification analysis
(WCRQA; Webber Jr \& Zbilut, 2005), we analyzed the co-evolution of social
cohesion online and real-world events. Specifically, using heterogeneous
data from March 30, 2012, to June 15, 2012, we compared the social cohesion
of Syrian tweets with the counts
of all, positive, and negative real-world events.

We hypothesized that the count of all events and the count of negative
events would show high levels of determinism, as this time period was
marked by daily violence which often leads to increased salience. We also
hypothesized that the count of positive events would not show patterns of
determinism, as these events were likely overshadowed by the salient
negative daily occurrences.
After testing these hypotheses, we then conducted exploratory analyses
to further examine the patterns identified with CRQA. Using WCRQA,
we sought to identify shifts in the
dynamics of the systems that could then be linked to real-world shifts in
conflict frequency.
Through these analyses, our primary goal in the current work was to uncover
(1) how different intensities during times of turmoil may enhance or deter
social cohesion in social media communication and (2) how social mobilization
may be evident in the dynamic coupling of cohesion and real-world events.

@@

The social cohesion or linguistic similarity of daily tweets was calculated
by quantifying the frequency of shared content words with respect to those
words' frequencies in the entire tweet corpus. If a linguistic term has a
high frequency in one tweet but has a low frequency in the whole tweet
corpus, then it is more likely a content word and is this given more weight
in calculating social cohesion.

The tweets were first sorted into ascending order by timestamp. They were
then grouped into successive sets of five tweets. Inverse frequency
weighting for the overlap words and all words was conducted for each pair
of two tweets in the group of five, then divided to get the social cohesion
score for that pair (see Eq. 1-3). This resulted in ten social cohesion
values for each group of five tweets, which were then averaged. Finally,
the mean of all average cohesion values for all groups was taken to get
a social cohesion score for each day.

@@

The social cohesion calculation was performed on a subset of the corpus
containing (recalculate with new date span) English and (recalculate with
new date span) Arabic tweets. All hashtags and URLs were included in the
analysis. This method is similar to that of inverse document frequency (IDF).
For more information on IDF, see Jones (1972) and
Robertson (2004).

@@

Using the Integrated Crisis Early Warning System (ICEWS) developed at the
Defense Advanced Research Project Agency, the daily events in Syria were
scored with an intensity rating ranging from -10 to 10. Negative events,
such as riots and attacks, are coded with low negative numbers while
positive, peaceful events are coded with high positive numbers. This
well-validated coding scheme allows for numerical quantification of
different event types.

For both the target only and the target and source data, time series of
the count of all events were generated by summing the total number of events
on each day, irrespective of ICEWS score. Similarly, time series of count
of positive events were generated by summing the number of ICEWS events
for each day that had an intensity rating greater than zero. Lastly,
time series of count of negative events were generated by summing the number
of ICEWS events for each day that had an intensity rating less than zero.

@@

We first used cross-recurrence quantification analysis
(CRQA; Zbilut et al., 1998) to investigate the temporal patterns of
social cohesion and real-world events in Syria during Arab Spring. As
an extension of RQA, which captures the patterns of revisited states of
a single time series (Webber Jr \& Zbilut, 1994; Zbilut \& Webber Jr, 1992), CRQA
allows for two time series to be projected onto one another to investigate
shared states and trajectories. Nonlinear analyses such as CRQA can reveal
dynamics that are not captured by traditional linear methods.

In order to perform CRQA, the time series need to be converted to the same
scale, while still preserving a sampling rate that respects the native
timescale of the system
(cf.~Nyquist frequency in signal processing; Grenander, 1959).
We converted each time series (e.g., Figure \ref{fig:raw-ts}) into deciles
(e.g., Figure \ref{fig:deciled-ts}) to analyze the dynamics of different
levels of social cohesion and counts of all, positive, and negative events
on each day. In CRQA, a recurrent point is when two signals share the same
state. These shared states can be at the same time or at later points in
the time series.

@@

A recurrence plot (RP) is the visual representation of CRQA. To construct
a recurrence plot, two time series are plotted against each other on the
X and Y axes, and at every point where they share the same state, and point
in placed on the plot to indicate recurrence. Visually analyzing these plots
can help reveal patterns of structure and recurrence. The line y = x is known
as the line of identity. This line falls where the time series match
temporally, and a filled line would indicate perfect synchrony between the
two time series. See
Figures \ref{fig:plot-rp-targ-all}-\ref{fig:plot-rp-source-targ-neg}
for examples of RPs.

@@

The percentage of recurrent points to total possible points is a metric
called recurrence rate (RR). This metrics gives a general description of
how frequently the time series are sharing the same states. In the current
study, a recurrent point would indicate that the level of social cohesion
and frequency of events were at the same intensity.

Recurrent points that occur in successive moments form line structures.
These are conceptualized as shared trajectories, where the two time series
are moving together across time in a shared state. The percentage of
recurrent points that make up diagonal line structures (i.e., two or more
consecutive points) is a metric called percent determinism (DET). For the
current project, a high DET would indicate that not only are the levels of
social cohesion and frequency of events occurring at similar intensities,
but they are doing so across consecutive days.

The average length of the line structures is called mean line (L). This can
be conceptualized as the average amount of time that the two time series move
together before one or both shift states. The longest line length is called
max line (maxL). maxL indicates the longest instance of being \enquote{stuck} in a
particular state together. The total number of lines is called NRLINE. The
more lines there are, the more coordinated the system may be. The
complexity or variability of line lengths is captured by a metric called
entropy (ENTR) and entropy while accounting for the total number of lines in
the plot is normalized entropy (rENTR). These entropy metrics help indicate
the stability and predictability of the system. When one of the signals gets
\enquote{stuck} in a state and the other moves on to other states, this results in
vertical line structures. The proportion of points forming vertical line
structures is a metric called laminarity (LAM) and the average length of the
vertical trajectories is called trapping time (TT). For more detail on CRQA
metrics, see Coco and Dale (2018).

@@

To conduct the analyses on the decile data, the parameters were set to those
typically used in categorical CRQA. Delay was set to zero, embedding dimension
was set to 1, and radius was set to .001 to allow for only exact recurrent
matches. No normalization was done, and lines were considered two or more
consecutive recurrent points. The Theiler window was set to zero to include
points along the line of identity. For more details on the CRQA parameters and
how they impact the analysis, see Coco and Dale (2018).

@@

Using the deciled, real time series of social cohesion, count of all events,
count of positive events, and count of negative events, a permutation test was
done to determine statistical difference from chance for each CRQA metric.
Using the sample function in base R (R Core Team, 2019), 1000 permuted time series were
created for each of the real time series. A permutation test is similar to
bootstrapping, except that the sampling of observations is done without
replacement. CRQA was then run 1000 times for social cohesion and count of all
events, social cohesion and count of positive events, and social cohesion and
count of negative events using the permuted time series. This resulted in
1000 results for each of the CRQA metrics for each combination of time series.
Statistical significance from chance was set at the 95th and 99th percentiles
of the metrics from the permuted CRQ analyses. Given that a permutation test
is similar to bootstrapping but does not include replacement, all points were
included in each CRQA analysis, leaving RR the same across all analyses. Thus,
RR is not included in the inferential tests. For more information on permutation
tests, see Good (2013).

@@

We used windowed cross-recurrence quantification analysis
(WCRQA; Webber Jr \& Zbilut, 2005) to further investigate the evolution of the
temporal dynamics of social cohesion and real-world events. In this analysis,
a smaller \enquote{window} of time points is used to calculate CRQA and capture more
fine-grain dynamics of two time series. The window remains the same size and
\enquote{slides} diagonally up the line of identity. The changes in the CRQA metrics
across the windows can reveal changes in dynamics that can then be linked to
external, real-world events.

@@

Conducting WCRQA results in CRQA metrics (e.g., RR, DET, ENTR) for each
respective window. These metrics can then be plotted as a time series to
visualize their evolution over time. For these analyses, we decided to
visualize recurrence rate and determinism.

@@

In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, a window size must be decided on. The window size
will determine how many observations are used in each CRQA analysis. We
conducted WCRQA with a window size of 14 to investigate the dynamics over
two weeks and allow for the capture of more structure and thus a more
meaningful interpretation of DET. The second parameter to be chosen is a
step size. Window step size is how far the window moves up the line of identity
before conducting CRQA again. We chose 1 for the step size to understand
how the metrics evolve from day to day. All other CRQA metrics remained the
same as previously noted.

@@

To visualize the differences in the WCRQA metrics from those obtained by
chance, a permutation test was done to set upper and lower bounds for the
95th and 99th percentiles. Using a sample size of 14, one-thousand permuted
time series were generated without replacement for each real time series. The
upper and lower 95th and 99th percentiles were set based on the distribution
of CRQA metrics from these one-thousand CRQ analyses.

@@

CRQA was performed using the crqa package (Coco \& Dale, 2018) for R (R Core Team, 2019). We
conducted a permutation test to check for statistical difference from chance
for all analyses and metrics.

@@

A permutation test for the CRQA of social cohesion and count of all events
found a statistical difference from chance for DET, NRLINE, maxL, and LAM.
All other metrics were non-significant (Figure \ref{fig:plot-rp-targ-all};
\autoref{table-1}). A permutation test for the CRQA of social cohesion and
count of positive events found no statistical difference from chance for any
of the metrics other than TT (Figure \ref{fig:plot-rp-targ-pos};
\autoref{table-1}). A permutation test for the CRQA of social cohesion and
count of negative events found a statistical difference from chance for DET,
NRLINE, L, ENTR, LAM, and TT. All other metrics were non-significant (Figure
\ref{fig:plot-rp-targ-neg}; \autoref{table-1}). Looking at the recurrence
plots there appears, to be a shift in dynamics around the 36th observation
which corresponds to May 4th, 2012.

@@

A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, and LAM. All other metrics were non-significant (Figure \ref{fig:plot-rp-source-targ-all}; \autoref{table-2}). A permutation test for the CRQA of social cohesion and count of positive events found a statistical difference from chance for maxL, ENTR, and LAM (Figure \ref{fig:plot-rp-source-targ-pos}; \autoref{table-2}). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, and rENTR. All other metrics were non-significant (Figure \ref{fig:plot-rp-source-targ-neg}; \autoref{table-2}). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

@@

WCRQA was performed using the crqa package (Coco \& Dale, 2018) for R (R Core Team, 2019). Permutation tests
were conducted for each WCRQA to establish upper and lower 95th and 99th percentiles. In
the plots (see Figures \ref{fig:plot-wcrqa-targ-all} and \ref{fig:plot-wcrqa-source-targ-all}),
the 95th percentiles are represented by horizontal red lines and the 99th percentiles are
represented by horizontal orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there is a consistent period
of above chance recurrence rate from the 7th-12th and 36th-47th window
(Figure \ref{fig:plot-wcrqa-targ-all}). Determinism is largely above chance
from the 46th to the 61st window (Figure \ref{fig:plot-wcrqa-targ-all}). For
social cohesion and count of positive events there is a downward trend in RR
as the window moves with largly above chance RR from the 1st-18th and
32nd-45th windows. Determinism was at or above chance for the 8th-17th,
32nd-34th, and the 60th window (Figure \ref{fig:plot-wcrqa-targ-all}).
For social cohesion and count of negative events, recurrence rate stays above
or almost above chance from the 4th-7th and 43rd-47th windows, and determinism
is above or nearly above chance in the 1st and 2nd windows, with short intermittent
peaks in later windows (Figure \ref{fig:plot-wcrqa-targ-all}).

@@

For social cohesion and count of all events, there is a consistent period of
above or nearly above chance determinism from the 7th-13th and 26th-47th
windows (Figure \ref{fig:plot-wcrqa-source-targ-all}). Determinism also
shows above or nearly above chance values in windows 7th-17th and short,
intermittent peaks between the 34th and 61st windows (Figure
\ref{fig:plot-wcrqa-source-targ-all}). For social cohesion and count of
positive events, there is a downward trend in RR as the window moves with
largely above chance values in the first 20 windows and the 32nd-45th windows.
Determinism stays largely within chance with the exception of windows 13
through 19 (Figure \ref{fig:plot-wcrqa-source-targ-all}). For social cohesion
and negative events, there appears to be an extended time period where RR
(32nd-47th windows) and DET (35th-39th windows) were above of nearly above
chance (Figure \ref{fig:plot-wcrqa-source-targ-all}). These windows span from
May 3rd, 2012 to May 14th, 2012.

@@

In the present study, we investigated the dynamics of real-world events
and online communication on Twitter in Syria during Arab Spring. We examined
how social mobilization both virtually and physically evolve over time in
tandem. Prior to now, the extent to which virtual mobilization on social
media platforms such as Twitter has been largely unexplored. Using the
nonlinear methods cross-recurrence quantification analysis and windowed
cross-recurrence quantification analysis, we investigated the theory that
virtual mobilization is increasingly reflective of real-world events
during times of strife.

@@

The significant result of DET for social cohesion and both the all events
and negative events time series confirms our hypothesis, that increased
salience of events is reflected in social cohesion on Twitter. The frequency
of all events and of negative events thus result in a significant number
of shared trajectories that are statistically different from chance. The
non-significant results for all metrics for positive event count also
confirm our hypothesis, and gives support to the theory that positive events
are less coupled to social cohesion than the more salient, negative events
during times of turmoil. The inevitable increased salience of event
frequency and event violence thus leads to increased coupling with Twitter
social cohesion.

@@

By adding Syria as a source in the data filtering, the dataset increased
in size by 1,690 samples. These additional observations are times where
Syria initiated an event with another country. Events initiated by Syria
but directed internally within the country were already captured by the
target only filtering. The increase in coupling between positive events
and social cohesion is thus due to the inclusion of these externally
directed events. Though CRQA metrics for count of all and count of
negative events with social cohesion were largely unaffected, count of
positive events and social cohesion showed significant levels of
determinism, number of lines, max line length, and laminarity that were
not present in the target filtered data. Given that the Syria during
this time was still under Authoritarian rule, the government may have
been initiating more positive external events to prevent intervention
by other countries or try to cover up what was happening internally.
We believe that the increased coupling of positive events is thus not
reflective of actual social cohesion, but happened by chance. The
inflation of the number of positive events when including source
observations is causing increased coupling, but it is more likely that
the Twitter users were using the platform to raise awareness of the
violent, internal events that were being inflicted on them by their
government officials and armies and not reporting on the positive
external events that did not directly affect them or their safety.
Further research should explore the semantic content of this tweet
corpus to determine what the socially cohered language was focused on.
By visual inspection of the recurrence plots (RPs;
Figures \ref{fig:plot-rp-targ-all}-\ref{fig:plot-rp-source-targ-neg}),
there appears to be a shift in the dynamics of these systems around
observation 35, or May 3rd, 2012. To further investigate these local
dynamics, we performed a sliding window technique known as windowed
cross-recurrence quantification analysis (WCRQA).

@@

The time period of differing dynamics, from observation 38 through 43,
appears to be captured by WCRQA of social cohesion and count of negative
events for both RR and DET (Figure \ref{fig:plot-wcrqa-targ-all}). The
amount of recurrence and determinism for these periods almost consistently
lie above chance, suggesting that the intensity of negative events coincide
with the degree of social cohesion found on Twitter, and that these states
carry across multiple days in a way that is different from that of chance.

The above-chance levels of determinism and recurrence rate are demonstrating
the coupling of social cohesion and counts of all and negative events during
times of strife. We can see this string of highly violent events reflected
in the mode ICEWS event for each day (Figure \ref{fig:plot-mode-event}),
where from May 1st, 2020 to May 15th, 2020, the most frequently occurring
event was almost consistently -10, the most violent event in the coding
system. These dates directly correlate with the above chance RR and DET
metrics for the WCRQA of social cohesion and count of negative events.
Furthermore, for WCRQA of social cohesion and positive event count during
this time period, DET was entirely within chance (Figure
\ref{fig:plot-wcrqa-targ-all}). This supports our hypothesis, as the
positive events are seemingly not coupled with the social cohesion on Twitter
during civil uprisings during the Arab Spring.

@@

WCRQA of recurrence rate on the source and target filtered data remained
largely the same for count of all events and social cohesion, suggesting
that large numbers on a given day will be reflective of increased levels
of social cohesion. Determinism showed more frequent above-chance metrics
for windows 39 and up, suggesting that the inclusion of externally
directed events lead to higher event counts during these days and then
was more reflective of the level of social cohesion on Twitter. WCRQA for
count of positive events and social cohesion remained largely the same for
RR and DET as the source only data. This again seems to support the idea
that these two aren't actually coupled, but rather that the inflation of
the positive counts lead to a spurious relationship. Count of negative
events and social cohesion also showed similar trends for RR and DET,
again attesting to the theory that negative events were coupled to social
cohesion on Twitter while positive events happened by chance.

@@

>>>>

Using cross-recurrence quantification analysis (CRQA; Zbilut et al., 1998)
and windowed cross-recurrence quantification analysis
(WCRQA; Webber Jr \& Zbilut, 2005), we analyzed the co-evolution of social
cohesion online and real-world events. Specifically, using heterogeneous
data from March 30, 2012, to June 15, 2012, we compared the social cohesion
of Syrian tweets with the counts
of all, positive, and negative real-world events derived from gold-standard
international event data (i.e., the Integrated Crisis Early Warning
System; Boschee et al., 2015).

We hypothesized that the count of all events and the count of negative
events would show high levels of determinism, as this time period was
marked by daily violence which often leads to increased salience. We also
hypothesized that the count of positive events would not show patterns of
determinism, as these events were likely overshadowed by the salient
negative daily occurrences.
After testing these hypotheses, we then conducted exploratory analyses
to further examine the patterns identified with CRQA. Using WCRQA,
we sought to identify shifts in the
dynamics of the systems that could then be linked to real-world shifts in
conflict frequency.
Through these analyses, our two primary goals in the current work were to uncover
(1) how different intensities during times of turmoil may enhance or deter
social cohesion in social media communication and (2) how social mobilization
may be evident in the dynamic coupling of cohesion and real-world events.

@@

The social cohesion or linguistic similarity of daily tweets was calculated
by quantifying the frequency of shared content words with respect to those
words' frequencies in the entire tweet corpus. If a linguistic term has a
high frequency in one tweet but has a low frequency in the whole tweet
corpus, then it is more likely a content word and is thus given more weight
in calculating social cohesion.
This method is similar to inverse document frequency (IDF).
For more information on IDF, see Jones (1972) and
Robertson (2004).

The tweets were first sorted into ascending order by timestamp. They were
then grouped into successive sets of five tweets. Inverse frequency
weighting for the overlap words and all words was conducted for each pair
of two tweets in the group of five, then divided to get the social cohesion
score for that pair (see Eq. 1-3). This resulted in ten social cohesion
values for each group of five tweets; these values were then averaged to get
a social cohesion score for the entire set of five tweets. Finally,
the mean of all average cohesion values for all groups was taken to get
a social cohesion score for each day.

@@

The social cohesion calculation was performed on the tweet corpus
containing (recalculate with new date span) English and (recalculate with
new date span) Arabic tweets. All hashtags and URLs were included in the
analysis.

@@

Using the Integrated Crisis Early Warning System (ICEWS) developed at the
Defense Advanced Research Project Agency, the daily events were
scored with an intensity rating ranging from -10 to 10. Negative
events---such as riots and attacks---are coded with low negative numbers, while
positive or peaceful events are coded with high positive numbers. This
well-validated coding scheme allows for numerical quantification of
different event types.

For both the target only and the target and source data described above, time series of
the count of all events were generated by summing the total number of events
on each day, irrespective of ICEWS score. Similarly, time series of count
of positive events were generated by summing the number of ICEWS events
for each day that had an intensity rating greater than zero. Lastly,
time series of count of negative events were generated by summing the number
of ICEWS events for each day that had an intensity rating less than zero.

@@

We first used cross-recurrence quantification analysis
(CRQA; Zbilut et al., 1998) to investigate the temporal patterns of
social cohesion and real-world events in Syria during Arab Spring. CRQA
is an extension of RQA, which captures the patterns of revisited states of
a single time series (Webber Jr \& Zbilut, 1994; Zbilut \& Webber Jr, 1992). CRQA
allows for two time series to be projected onto one another to investigate
shared states and trajectories.
In CRQA, we track these shared states by identifying \emph{recurrent points},
or intersections of time during which two signals shared or revisited the same
state. These shared states can be at the same time or at later points in
the time series. We can further chart shared trajectories by examining
the occurrences and patterns of multiple sequential recurrent points.

To perform CRQA, the time series must be converted to the same
scale, while still preserving a sampling rate that respects the native
timescale of the system
(cf.~Nyquist frequency in signal processing; Grenander, 1959).
Because the social cohesion and event data have different scales and values,
we converted each time series (e.g., Figure \ref{fig:raw-ts}) into deciles
(e.g., Figure \ref{fig:deciled-ts}) to analyze the relative dynamics of different
levels of social cohesion and counts of all, positive, and negative events
on each day.
We implemented CRQA using the \texttt{crqa} package (Coco \& Dale, 2018) for R (R Core Team, 2019).

@@

A cross-recurrence plot (CRP) is the visual representation of CRQA. To construct
a cross-recurrence plot, two time series are plotted against each other on the
\(x\) and \(y\) axes. At every point where they share the same state, a point
in placed on the plot to indicate recurrence. (For categorical CRQA, this process
is similar to the Cartesian product.)

We can visually analyze these plots
to reveal patterns of structure and recurrence. The line \(y = x\) is known
as the \emph{line of identity}. This line falls where the time series match
temporally. A filled line would indicate perfect in-phase synchrony between the
two time series. See
Figures \ref{fig:plot-rp-targ-all}-\ref{fig:plot-rp-source-targ-neg}
for examples of RPs.

@@

While visual inspections can be informative, we can derive a variety of
quantitative metrics from the cross-recurrence plot. Each of these metrics
provides information into the pattern and dynamics of the new shared system
between the original time series. These metrics are also critical for inferential
statistics. We will focus most directly on two of these metrics:
\emph{recurrence rate} and \emph{determinism}.

The percentage of recurrent points to total possible points is a metric
called the \emph{recurrence rate} (\emph{RR}). This metrics gives a general description of
how frequently the time series are sharing the same states. In the current
study, a recurrent point would indicate that the level of social cohesion
and frequency of events were at the same intensity.

Recurrent points that occur in successive moments form line structures.
These are conceptualized as shared trajectories, where the two time series
are moving together across time in a shared state. The percentage of
recurrent points on the plot that fall along diagonal line structures
(i.e., two or more
consecutive points) is a metric called \emph{percent determinism}
(or simply \emph{determinism}; \emph{DET}). For the
current project, a high DET would indicate that not only are the levels of
social cohesion and frequency of events occurring at similar intensities,
but they are doing so across consecutive days.

A number of other metrics provide insight not only into the amount of structure
between the systems but about the nature of this shared structure.
The average length of the line structures is called \emph{mean line} (L), which can
be conceptualized as the average amount of time that the two time series move
together before one or both shift states. The longest line length is called
\emph{max line} (\emph{maxL}); it indicates the longest instance of being \enquote{stuck} in a
particular state together and is a measure of the stability of the system.
The total \emph{number of lines} on the plot (\emph{NRLINE})
scales with structure: The more lines that exist, the more coupled the systems are. The
complexity or variability of line lengths is captured by a metric called
\emph{entropy} (\emph{ENTR})---which may also be normalized by the total number of lines
(i.e., \emph{normalized entropy}; \emph{rENTR})---and indicates
the stability and predictability of the system. When one of the signals gets
\enquote{stuck} in a state and the other moves on to other states, this results in
vertical line structures, which are quantified in \emph{laminarity} (\emph{LAM};
i.e., the proportion of points forming vertical line structures)
and trapping time (\emph{TT}; i.e., the average length of the
vertical trajectories). For more detail on CRQA metrics, see Coco and Dale (2018).

@@

To conduct the analyses on the decile data, the parameters were set to those
typically used in categorical CRQA: Delay was set to \(0\), embedding dimension
was set to \(1\), and radius was set to \(.001\) to allow for only exact recurrent
matches. No normalization was done, and lines were considered two or more
consecutive recurrent points. The Theiler window was set to \(0\) to include
points along the line of identity. For more details on the CRQA parameters and
how they impact the analysis, see Coco and Dale (2018).

@@

Unlike some nonlinear analyses (e.g., fractal analyses), CRQA metrics do not have
inherently meaningful values. As a result, CRQA is best used as a relative
metric---that is, comparing the CRQA values across conditions or subsets
of the data. Typically, meaningful differences between values are determined with
traditional inferential statistics. While this may not be a problem for
experimentally derived datasets, it can pose a barrier to researchers using
small-\(n\) data or conducting data-informed case studies.

For CRQA, this can be addressed using \emph{permutation tests}.
Similar to more established methods of statistical significance in CRQA research
(which create surrogate time series; e.g., Paxton \& Dale, 2017),
the permutation test approach to significance testing allows the researcher
to use a time series as its own baseline by testing whether the \emph{structure}
of the unfolding events of the time series---rather than the \emph{raw frequency} of the
events within the time series---cohere together more than would be expected
by chance. A permutation test is similar to
bootstrapping, except that the sampling of observations is done without
replacement. In this approach, we simply create a large number of permutations
of the original time series (i.e., breaking up the dependency across time points but
preserving the frequencies of the events), conduct CRQA for each of those
permutations, and then examine whether the real time series' CRQA metrics
exceed those that we would expect to see by chance. The proportion of times that
the real time series' values exceeds the surrogate time series' values can
be interpreted as the alpha criterion to determine significance.
For more information on permutation
tests, see Good (2013).

Using the real deciles of social cohesion, count of all events,
count of positive events, and count of negative events, we conducted a series
of permutation tests
(one for the comparison between social cohesion and each type of event count)
to determine statistical difference from chance for each CRQA metric.
Using the \texttt{sample} function in base R (R Core Team, 2019), \(1000\) permuted time series were
created for each of the real time series.\\
CRQA was then run \(1000\) times for social cohesion and count of all
events, social cohesion and count of positive events, and social cohesion and
count of negative events using the permuted time series. This resulted in
\(1000\) results for each of the CRQA metrics for each combination of time series.
Statistical significance from chance was set at the 95th and 99th percentiles
of the metrics from the permuted CRQ analyses. Given that a permutation test
is similar to bootstrapping but does not include replacement, all points were
included in each CRQA analysis, leaving RR constant across all permutations. Thus,
RR is not included in the inferential tests.

@@

We used windowed cross-recurrence quantification analysis
(WCRQA; Webber Jr \& Zbilut, 2005) to further investigate the evolution of the
temporal dynamics of social cohesion and real-world events. In this analysis,
a smaller \enquote{window} of time points is used to calculate CRQA and capture more
fine-grained dynamics of two time series. The window remains the same size and
\enquote{slides} diagonally up the line of identity. The changes in the CRQA metrics
across the windows can reveal changes in dynamics that can then be linked to
external, real-world events.
We implemented WCRQA using the \texttt{crqa} package (Coco \& Dale, 2018) for R (R Core Team, 2019).

@@

Conducting WCRQA results in CRQA metrics (e.g., RR, DET, ENTR) for each
respective window. These metrics can then be plotted as a time series to
visualize their evolution over time. For these analyses, we decided to
focus on recurrence rate and determinism.

@@

In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, the \emph{window size}
will determine how many observations are used in each CRQA analysis. We
conducted WCRQA with a window size of \(14\) to investigate the dynamics over
two weeks and allow for the capture of more structure and thus a more
meaningful interpretation of DET.

The second parameter to be chosen is a \emph{window step size}.
Window step size is how far the window moves up the line of identity
before conducting CRQA again. We chose \(1\) for the step size to understand
how the metrics evolve from day to day.

All other CRQA metrics remained the same as those previously noted.

@@

To visualize the differences in the WCRQA metrics from those obtained by
chance, a permutation test was done to set upper and lower bounds for the
95th and 99th percentiles. Using a sample size of 14, \(1000\) permuted
time series were generated without replacement for each real time series. The
upper and lower 95th and 99th percentiles were set based on the distribution
of the permuted WCRQA metrics.

@@

@@

A permutation test for the CRQA of social cohesion and count of all events
found a statistical difference from chance for DET, NRLINE, maxL, and LAM.
All other metrics did not reach significance (Figure \ref{fig:plot-rp-targ-all};
\autoref{table-1}).

A permutation test for the CRQA of social cohesion and
count of positive events found no statistically significant difference for any
of the metrics other than TT (Figure \ref{fig:plot-rp-targ-pos};
\autoref{table-1}).

A permutation test for the CRQA of social cohesion and
count of negative events found a statistically significant difference for DET,
NRLINE, L, ENTR, LAM, and TT. All other metrics did not reach significance
(Figure \ref{fig:plot-rp-targ-neg}; \autoref{table-1}).

A qualitative investigation of the recurrence plots suggests
a shift in dynamics around the 36th observation,
which corresponds to May 4th, 2012.

@@

A permutation test for the CRQA of social cohesion and count of all
events found a statistical difference from chance for DET, NRLINE, and
LAM. All other metrics were non-significant (Figure
\ref{fig:plot-rp-source-targ-all}; \autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of positive
events found a statistical difference from chance for maxL, ENTR, and LAM
(Figure \ref{fig:plot-rp-source-targ-pos}; \autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of negative
events found a statistical difference from chance for DET, NRLINE, and rENTR.
All other metrics were non-significant (Figure
\ref{fig:plot-rp-source-targ-neg}; \autoref{table-2}).

Looking qualitatively at the recurrence plots, there again appeared
to be a shift in dynamics around the 36th observation (May 4th, 2012).

@@

In the plots (see Figures \ref{fig:plot-wcrqa-targ-all} and
\ref{fig:plot-wcrqa-source-targ-all}), the 95th percentiles are represented
by horizontal red lines, and the 99th percentiles are
represented by horizontal orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there was a consistent period
of above-chance RR from the 7th-12th and 36th-47th window
(Figure \ref{fig:plot-wcrqa-targ-all}). DET was largely above chance
from the 46th to the 61st window (Figure \ref{fig:plot-wcrqa-targ-all}).

For social cohesion and count of positive events, there was a downward trend in RR
as the window moves with largely above-chance RR from the 1st-18th and
32nd-45th windows. DET was at or above chance for the 8th-17th,
32nd-34th and at the 60th window (Figure \ref{fig:plot-wcrqa-targ-all}).

For social cohesion and count of negative events, RR stayed above
or almost above chance from the 4th-7th and 43rd-47th windows. DET
was above or nearly above chance in the 1st and 2nd windows, with intermittent
peaks in later windows (Figure \ref{fig:plot-wcrqa-targ-all}).

@@

For social cohesion and count of all events, there was a consistent period of
above- or nearly above-chance DET from the 7th-13th and 26th-47th
windows (Figure \ref{fig:plot-wcrqa-source-targ-all}). DET also
showed above- or nearly above-chance values in windows 7th-17th and
intermittent peaks between the 34th and 61st windows (Figure
\ref{fig:plot-wcrqa-source-targ-all}).

For social cohesion and count of
positive events, there was a downward trend in RR as the window moved, with
largely above-chance values in the first 20 windows and the 32nd-45th windows.
DET stayed largely within chance, with the exception of windows 13
through 19 (Figure \ref{fig:plot-wcrqa-source-targ-all}).

For social cohesion
and negative events, there was an extended period of time during which RR
(32nd-47th windows) and DET (35th-39th windows) were above or nearly above
chance (Figure \ref{fig:plot-wcrqa-source-targ-all}). These windows span from
May 3rd--14th, 2012.

@@

In the present study, we investigated the dynamics of real-world events
and online communication on Twitter in Syria during Arab Spring. We examined
how social mobilization both virtually and physically evolve over time in
tandem. Using the
nonlinear dynamical time series analyses, we investigated the theory that
virtual mobilization is increasingly reflective of real-world events
during times of strife.

@@

The significant result of DET for social cohesion with all and negative
events supports our hypothesis that
event salience is coupled with online social cohesion The frequency
of all and negative events are thus associated with greater
shared trajectories with social cohesion that would be expected by chance.
However, when we specifically target positive events, the connection
to social cohesion fades, providing further
support for our hypothesis. Positive events
may be less coupled to social cohesion than more salient negative events
during times of turmoil; by contrast, overall event frequency and
negative event frequency are coupled with Twitter social cohesion.

@@

By adding Syria as a source in the data filtering, the dataset increased
in size by 1,690 samples. These additional observations are times where
Syria initiated an event with another country. Events initiated by Syria
but directed internally within the country were already captured by the
target-only filtering. The increase in coupling between positive events
and social cohesion can thus only be due to the externally
directed events. Though the coupling between social cohesion with
negative and all events were largely unchanged,
positive events and social cohesion showed greater stability
in their coupling (reflected in maxL); while TT was no longer
significant in these analyses, LAM (related to TT) did
reach statistical significance.

While interesting, the underlying meaning or drivers of this change
in results requires further investigation that incorporates the
actual content of the tweets themselves. One possible reason (among
many) may be that the finding was spurious. The positive external
events from Syria may have been self-serving, aiming to prevent intervention
by other countries or to cover up what was happening internally.
During these moments, the increased cohesion may reflect a surge from
the Syrian people trying to raise awareness of the situation or
to decry the Syrian government's attempts to improve their international
standing. In this case, the increased coupling of positive events would
coupled not with shared positive affect in social media but instead
to shared negative affect in social media in response. However, we
cannot test this hypothesis with the current data; future\\
research should explore the semantic content of this tweet
corpus to determine the nature of the socially cohered language.

@@

The time period of differing dynamics---from observation 38 through
43---appears to be captured by the RR and DET values in
WCRQA of social cohesion and count of negative
events (Figure \ref{fig:plot-wcrqa-targ-all}). The
amount of recurrence and determinism for these periods almost consistently
lie above chance, suggesting that the intensity of negative events coincide
with the degree of social cohesion found on Twitter and that these
shared trajectories extend across multiple days.

The above-chance levels of determinism and recurrence rate demonstrate
the coupling of social cohesion to both total and negative events during
times of strife. During this period,
the modal ICEWS event for each day (Figure \ref{fig:plot-mode-event})
during May 1st---15th was almost consistently -10, the most violent
event category. These dates directly correlate with the above-chance RR and DET
metrics for the WCRQA of social cohesion and count of negative events.
Furthermore, for WCRQA of social cohesion and positive event count during
this time period, DET was entirely within chance (Figure
\ref{fig:plot-wcrqa-targ-all}). This supports our hypothesis, as the
positive events are seemingly not coupled with the social cohesion on Twitter
during civil uprisings during the Arab Spring.

@@

WCRQA of recurrence rate on the source and target filtered data remained
largely the same for count of all events and social cohesion, suggesting
that large numbers on a given day will be reflective of increased levels
of social cohesion. DET showed more frequent above-chance metrics
for windows 39 and up, suggesting that the inclusion of externally
directed events lead to higher event counts during these days and then
was more reflective of the level of social cohesion on Twitter.

WCRQA metrics for positive-only and negative-only event counts with
social cohesion remained largely the same for
RR and DET as the source-only data. This may support the ideas suggested
in the CRQA section, but again, only future work including the \emph{content}
of those tweets can directly speak to that explanation.

@@

\leavevmode\hypertarget{ref-paxton2017interpersonal}{}%
Paxton, A., \& Dale, R. (2017). Interpersonal movement synchrony responds to high-and low-level conversational constraints. \emph{Frontiers in Psychology}, \emph{8}, 1135.
